---
title: "housing"
author: "ianalexmac"
date: "1/27/2022"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### library
```{r}
#install.packages('corrplot')
#install.packages('car')
#install.packages('mgcv')
install.packages('visreg')


library(corrplot)
library(car)
library(mgcv)
library(visreg)

```


```{r}
test <- read.csv("~/R/housing/test.csv")
train <- read.csv("~/R/housing/train.csv")

test$SalePrice <- 0
data <- rbind(train, test)

meta <- read.delim("~/R/housing/data_description.txt")


```


### convert character columns to factor,
number columns to integers
```{r}

data[sapply(data, is.character)] <- lapply(data[sapply(data, is.character)], as.factor)
data[sapply(data, is.numeric)] <- lapply(data[sapply(data, is.numeric)], as.integer)

test[sapply(test, is.character)] <- lapply(test[sapply(test, is.character)], as.factor)
test[sapply(test, is.numeric)] <- lapply(test[sapply(test, is.numeric)], as.integer)

train[sapply(train, is.character)] <- lapply(train[sapply(train, is.character)], as.factor)
train[sapply(train, is.numeric)] <- lapply(train[sapply(train, is.numeric)], as.integer)

summary(data)


#####tidyverse version:
###data <- data %>% mutate_if(is.factor, as.character)

```

### explore NAs
```{r}

any(is.na(data))

apply(data, 2, function(x) any(is.na(x)))

apply(data, 2, function(x) sum(is.na(x)))

```

### cleaning NAs
NA columns (* = easy resolution)
  LotFrontage   486
  Alley         2721
  MSZoning      4
  Utilities     2
  Exterior1st   1*
  Exterior2nd   1*
  MasVnrArea    23
  BsmtQual      81
  BsmtCond      82
  BsmtExposure  82
  BsmtFinType1  79
  BsmtFinSF1    1
  BsmtFinSF2    1
  BsmtUnfSF     1
  TotalBsmtSF   1
  Electrical    1*
  BsmtdataBath  2
  BsmtHalfBath  2
  KitchenQual   1
  Functional    2
  FireplaceQu   1420
  GarageType    157
  GarageYrBlt   159
  GarageFinish  159
  GarageCars    1
  GarageArea    1
  GarageQual    159*
  GarageCond    159*
  PoolQC        2909
  Fence         2348
  MiscFeature   2814  (105 total, 95 are Shed)
  SaleType      1*
  
  
oddballs:
  Neighborhood 1428 (Other)
  Condition1    35 (Other)
  Condition2    2  (Other)
  HouseStyle    27 (Other)
  RoofMatl      2  (Other)
  
probably drop:
  Alley
  PoolQC


### cut and remove from data
```{r}


data$Fence <- NULL
data$MiscFeature <- NULL
data$Alley <- NULL
data$PoolQC <- NULL
data$GarageType <- NULL
data$GarageYrBlt <- NULL
data$GarageCond <- NULL
data$GarageQual <- NULL
data$GarageFinish <- NULL
data$FireplaceQu <- NULL
data$MasVnrType <- NULL
data$MasVnrArea <- NULL
data$BsmtQual <- NULL
data$BsmtCond <- NULL
data$BsmtExposure <- NULL
data$BsmtFinType1 <- NULL
data$BsmtFinType2 <- NULL
data$LotFrontage <- NULL

```


### blatant assumptions
```{r}

data$MSZoning[is.na(data$MSZoning)] <- 'RL'
data$SaleType[is.na(data$SaleType)] <- 'WD'
data$Functional[is.na(data$Functional)] <- 'Typ'
data$KitchenQual[is.na(data$KitchenQual)] <- 'TA'
data$Exterior1st[is.na(data$Exterior1st)] <- 'VinylSd'
data$Exterior2nd[is.na(data$Exterior2nd)] <- 'VinylSd'
data$Utilities[is.na(data$Utilities)] <- 'AllPub'
data$BsmtdataBath[is.na(data$BsmtdataBath)] <- 0
data$BsmtFullBath[is.na(data$BsmtFullBath)] <- 0
data$BsmtHalfBath[is.na(data$BsmtHalfBath)] <- 0
data$BsmtFinSF1[is.na(data$BsmtFinSF1)] <- 0
data$BsmtFinSF2[is.na(data$BsmtFinSF2)] <- 0
data$BsmtUnfSF[is.na(data$BsmtUnfSF)] <- 0
data$TotalBsmtSF[is.na(data$TotalBsmtSF)] <- 0
data$GarageCars[is.na(data$GarageCars)] <- 0
data$GarageArea[is.na(data$GarageArea)] <- 0
data$Electrical[is.na(data$Electrical)] <- 'SBrkr'

any(is.na(data))

summary(data)
```


### split into train and test
```{r}
train <- data[1:1460,]
test <- data[1461:2919,]

any(is.na(test))
any(is.na(train))


```




### correlation matrix
```{r}
int.data <- sapply(data, is.integer)
corrplot(cor(data[,int.data]),type = "lower", mar = c(0,0,0,0))

```

### scatterplot matrix
```{r}
pairs(~YearBuilt + LotArea + GrLivArea + TotalBsmtSF + OverallQual, data = data)

```

### scatterplot exploration
```{r}
scatterplot(SalePrice ~ YearBuilt, data=train,  xlab="Year Built", ylab="Sale Price", grid=FALSE)

scatterplot(SalePrice ~ GrLivArea, data=train,  xlab="Square Footage", ylab="Sale Price", grid=FALSE)

```


### regression model
```{r}
#everthing!
lm_model_all <- lm(SalePrice ~ ., data=train)
summary(lm_model_all)

#more targeted 
lm_model_some <- lm(SalePrice ~ LotArea + OverallQual + OverallCond + YearBuilt + 
                    BsmtFinSF1 + BsmtFinSF2 + BsmtUnfSF + ExterQual + RoofMatl +
                    X1stFlrSF + X2ndFlrSF + BedroomAbvGr + KitchenQual + KitchenAbvGr, data=train)
summary(lm_model_some)
AIC(lm_model_some, lm_model_all)

gam_model_some <- gam(SalePrice ~ s(LotArea) + OverallQual + OverallCond + s(YearBuilt) + 
                    s(BsmtFinSF1) + s(BsmtFinSF2) + s(BsmtUnfSF) + ExterQual + RoofMatl +
                    s(X1stFlrSF) + s(X2ndFlrSF) + BedroomAbvGr + KitchenQual + KitchenAbvGr, data=train)
summary(gam_model_some)
visreg(gam_model_some)

#exp
gam_play <- gam(SalePrice ~ s(LotArea) +s(YearBuilt),
                data = train)
summary(gam_play)
visreg(gam_play)


AIC(lm_model_some, gam_model_some, gam_play)

```

